---
title: "3 - What makes a model? - Classwork"
subtitle: "Practical Machine Learning with tidymodels"
editor_options:
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Setup

Setup from deck 2:

```{r}
library(tidymodels)

set.seed(123)
cls_split <- initial_split(cls_data_2026, prop = 0.8)
cls_train <- training(cls_split)
cls_test <- testing(cls_split)
```

## Your turn

How do you fit a linear model in R?

How many different ways can you think of?

## To specify a model

```{r}
logistic_reg()
```

Engine - some models have a default engine, others don't

```{r}
logistic_reg() |>
  set_engine("glmnet")
```

```{r}
logistic_reg() |>
  set_engine("stan")
```

Mode - some models have a default mode, others don't

```{r}
decision_tree()
```

```{r}
decision_tree() |>
  set_mode("classification")
```

All available models are listed at <https://www.tidymodels.org/find/parsnip/>

## Your turn

Run the code below, then edit it to use a logistic regression model.

Extension/Challenge: Edit this code to use a different model. For example, try
using a conditional inference tree as implemented in the partykit package by
changing the engine - or try an entirely different model type!

```{r}
#| label: tree_spec

tree_spec <-
  decision_tree() |>
  set_mode("classification")

tree_spec

logistic_spec <-
  logistic_reg() |>
  set_engine("glmnet") |>
  set_mode("classification")
logistic_spec
```

## A model workflow

```{r}
tree_spec <-
  decision_tree() |>
  set_mode("classification")
```

Fit parsnip specification:

```{r}
tree_spec |>
  fit(class ~ ., data = cls_train)
```

Fit with a workflow:

```{r}
workflow() |>
  add_formula(class ~ .) |>
  add_model(tree_spec) |>
  fit(data = cls_train)
```

"Shortcut" by specifying the preprocessor and model spec directly in the
`workflow()` call:

```{r}
workflow(class ~ ., tree_spec) |>
  fit(data = cls_train)
```

## Your turn

Run the code below, then edit it to make a workflow with your own model of
choice.

Extension/Challenge: Other than formulas, what kinds of preprocessors are
supported?

```{r}
#| label: tree_wflow

tree_spec <- decision_tree() |>
  set_mode("classification")

tree_wflow <- workflow() |>
  add_formula(class ~ .) |>
  add_model(tree_spec)

tree_wflow

logistic_spec <- logistic_reg() |>
  set_mode("classification") |>
  set_engine("glm")

logistic_wflow <- workflow() |>
  add_formula(class ~ .) |>
  add_model(logistic_spec)

logistic_wflow
```

## Predict with your model

```{r}
tree_fit <-
  tree_wflow |>
  fit(data = cls_train)

logistic_fit <-
  logistic_wflow |>
  fit(data = cls_train)
```

## Your turn

Run:

```{r}
predict(tree_fit, new_data = cls_test)
# for logistic model
predict(logistic_fit, new_data = cls_test)
```

What do you notice about the structure of the result?

## Your turn

Run:

```{r}
augment(tree_fit, new_data = cls_test)

# for logistic model
augment(logistic_fit, new_data = cls_test)
```

How does the output compare to the output from `predict()`?

## Understand your model

```{r}
library(rpart.plot)
tree_fit |>
  extract_fit_engine() |>
  rpart.plot(roundint = FALSE)
```

## Your turn

Try extracting the model engine object from your fitted workflow!

```{r}
preds <- bind_rows(
  augment(logistic_fit, cls_test) |> mutate(model = "logistic"),
  augment(tree_fit, cls_test) |> mutate(model = "tree")
)

preds |>
  group_by(model) |>
  metrics(truth = class, estimate = .pred_class, .pred_class_1)
logistic_fit |>
  extract_fit_engine()

# on créé la courbe ROC des 2 modèles pour comparer
preds |>
  group_by(model) |>
  roc_curve(truth = class, .pred_class_1) |>
  autoplot()
```

Sur cls_test (n = 200), les deux modèles ont la même précision (accuracy = 0.895), mais logistic_fit domine sur les métriques probabilistes : ROC AUC de 0.964 contre 0.884 pour tree_fit, et une log loss plus faible (0.249 vs 0.343). La courbe ROC confirme visuellement que la régression logistique domine l'arbre sur presque toute la plage de seuils.

logistic_fit est donc le meilleur modèle pour ces données — il produit des probabilités mieux calibrées et discrimine mieux les deux classes, même si l'accuracy brute est identique.

Une mise en garde : ces comparaisons reposent sur un seul jeu de test — la variabilité d'échantillonnage n'est pas quantifiée ici (pas d'intervalle de confiance ni de rééchantillonnage).

What kind of object is it? What can you do with it?

⚠️ Never `predict()` with any extracted components!

You can also read the documentation for object extraction:
https://workflows.tidymodels.org/reference/extract-workflow.html
